---
title: "Data preparation and Exploratory 1"
output:
  pdf_document: default
  html_notebook: default
---


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).


---

# Load de data

```{r}
alldata<- read.csv("../data/train.csv")
alldata$dataset<-runif(nrow(alldata))


training <- alldata[alldata$dataset<=.6,]
validation <- alldata[alldata$dataset>.6 & alldata$dataset<=.8,]
testing <- alldata[alldata$dataset>.8,]


training$dataset <-NULL;
validation$dataset <-NULL;
testing$dataset  <-NULL;
alldata<-NULL;


str(training)
```

#Change type for cathegorical vars


```{r}
training$Survived<-factor(training$Survived)
training$Pclass<-factor(training$Pclass)
str(training)

```

```{r}
summary(training)
```

Plots
```{r}
plot(training$Survived)
```

```{r}
hist(training$Age)
```

```{r}
hist(training$SibSp)
```

```{r}
hist(training$Parch)
```

```{r}
hist(training$Fare)
```

```{r}
plot(training$Embarked)
```

# Bidimiensional distributions
Survival vs others

```{r}
with(training,plot(Pclass,Survived))
```


```{r}
with(training,plot(Sex,Survived))
```

```{r}
with(training,boxplot(Age,Survived))
```

```{r}
with(training,boxplot(SibSp,Survived))
```

```{r}
with(training,boxplot(Parch,Survived))
```

```{r}
with(training,boxplot(Fare,Survived))
```


```{r}
with(training,boxplot(log(Fare+.01),Survived))
```
```{r}
with(training,plot(Embarked,Survived))
```

# Check for data anomalies

No age?
```{r}

summary (training[is.na(training$Age),])
```

(no cabin for these people)

impostate the mean for these values 
```{r}
training[is.na(training$Age),]$Age<-mean(training$Age, na.rm = TRUE)
```

```{r}
summary(training$Age)

```



Null in "embarked"?

```{r}
training[training$Embarked=="",]
```

action -> remove these values

```{r}
train<-training[training$Embarked!="",]
```


possible outlier in Fare

```{r}
boxplot(training[training$Fare<500,]$Fare)


``` 

Eliminamos ese punto aunque la distrubicion sigue siendo muy sesgada

```{r}
train<-training[training$Fare<500,]

```


```{r}
summary(train)
```



# Prepared data: 

We discard artifial features and normalize the numerical features


```{r}
normalizationData <- list(
  ageMean=mean(training$Age),
  ageSD=sd(training$Age),
  SibSpMean=mean(training$SibSp),
  SibSpSD=sd(training$SibSp),
  ParchMean=mean(training$Parch),
  ParchSD=sd(training$Parch),
  FareMean=mean(training$Fare),
  FareSD=sd(training$Fare)
)

normalizationData
```


```{r}
training<-data.frame(
Pclass=training$Pclass,
Sex =training$Sex,
AgeNorm=(training$Age-normalizationData$ageMean)/normalizationData$ageSD,
SibSpNorm=(training$SibSp -normalizationData$SibSpMean)/normalizationData$SibSpSD,
ParchNorm=(training$Parch -normalizationData$ParchMean)/normalizationData$ParchSD,
FareNorm=(training$Fare -normalizationData$FareMean)/normalizationData$FareSD,
Embarked=training$Embarked,
Survived=training$Survived
  )
```

Check data cleaned

```{r}
summary(training)
```



save the clean data

```{r}
save(file = "../processed/training.dat", training)
save(file = "../processed/testing.dat", testing)
save(file = "../processed/validation.dat", validation)
```



