---
title: "Exploratory 1"
output:
  pdf_document: default
  html_notebook: default
---


Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).


---

# Load de data

```{r}
train<- read.csv("../data/train.csv")

str(train)
```

Change type for cathegorical vars


```{r}
train$Survived<-factor(train$Survived)
train$Pclass<-factor(train$Pclass)
str(train)

```

```{r}
summary(train)
```

Plots
```{r}
plot(train$Survived)
```

```{r}
hist(train$Age)
```

```{r}
hist(train$SibSp)
```

```{r}
hist(train$Parch)
```

```{r}
hist(train$Fare)
```

```{r}
plot(train$Embarked)
```

# Bidimiensional distributions
Survival vs others

```{r}
with(train,plot(Pclass,Survived))
```


```{r}
with(train,plot(Sex,Survived))
```

```{r}
with(train,boxplot(Age,Survived))
```

```{r}
with(train,boxplot(SibSp,Survived))
```

```{r}
with(train,boxplot(Parch,Survived))
```

```{r}
with(train,boxplot(Fare,Survived))
```


```{r}
with(train,boxplot(log(Fare+.01),Survived))
```
```{r}
with(train,plot(Embarked,Survived))
```

# Check for data anomalies

No age?
```{r}

summary (train[is.na(train$Age),])
```

(no cabin for these people)

impostate the mean for these values 
```{r}
train[is.na(train$Age),]$Age<-mean(train$Age, na.rm = TRUE)
```

```{r}
summary(train$Age)

```



Null in "embarked"?

```{r}
train[train$Embarked=="",]
```

action -> remove these values

```{r}
train<-train[train$Embarked!="",]
```


possible outlier in Fare

```{r}
boxplot(train[train$Fare<500,]$Fare)


``` 

Eliminamos ese punto aunque la distrubicion sigue siendo muy sesgada

```{r}
train<-train[train$Fare<500,]

```


```{r}
summary(train)
```



# Prepared data: 

We discard artifial features and normalize the numerical features


```{r}
normalizationData <- list(
  ageMean=mean(train$Age),
  ageSD=sd(train$Age),
  SibSpMean=mean(train$SibSp),
  SibSpSD=sd(train$SibSp),
  ParchMean=mean(train$Parch),
  ParchSD=sd(train$Parch),
  FareMean=mean(train$Fare),
  FareSD=sd(train$Fare)
)

normalizationData
```


```{r}
training<-data.frame(
Pclass=train$Pclass,
Sex =train$Sex,
AgeNorm=(train$Age-normalizationData$ageMean)/normalizationData$ageSD,
SibSpNorm=(train$SibSp -normalizationData$SibSpMean)/normalizationData$SibSpSD,
ParchNorm=(train$Parch -normalizationData$ParchMean)/normalizationData$ParchSD,
FareNorm=(train$Fare -normalizationData$FareMean)/normalizationData$FareSD,
Embarked=train$Embarked,
Survived=train$Survived
  )
```

Check data cleaned

```{r}
summary(training)
```



save the clean data

```{r}
save(file = "../data/cleantraining.dat", training)
```



